Home Nighttime Cough Monitoring

### 2025-07-01 Lilah Melzer/cough_monitor_analysis.qmd
cough monitor updates for Lilah

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
library(tidyverse)
library(ggstats)
library(corrplot)
library(patchwork)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

  labels = c("CF", "CF", "Non-CF", "Non-CF")

@@

df$Diagnosis <- relevel(df$Diagnosis, ref = "Non-CF")

@@

    cftr_mod, cftr_mod_type, mucolytic:antih

@@

```{r results='asis'}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
t1 <- tableby(Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
  nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + pre_fev_pp +
  total_sleep, data = demo, control = mycontrols)
summary(t1, pfootnote = T)
```

@@

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    .groups = "drop"
  )
# Exclude the people with 0 nights of data
per_person <- per_person %>% filter(`Nights With Data` != 0)
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

@@

```{r}
#| label: fig-coughs-dist
#| fig-cap: Histogram of Coughs Across All Nights
#| warning: false
hist <- ggplot(agg_counts, aes(x = Coughs)) +
  geom_histogram() +
  theme_bw() +
  theme(axis.title.y = element_blank())
hist
ggsave(paste0(home_dir, "/Dissemination/Figures/histogram.png"),
  plot = hist, width = 9, height = 6, units = "in"
)
```

@@

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
#| warning: false
fig2 <- ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
  x = Night, y = Coughs, group = factor(sid), color = Diagnosis
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.1) +
  geom_smooth(aes(group = Diagnosis)) +
  theme_bw()
fig2
ggsave(paste0(home_dir, "/Dissemination/Figures/fig2.png"),
  plot = fig2, width = 9, height = 6, units = "in"
)
```

>>>>

```{r setup}
#| include: false
library(Hmisc)
library(redcapAPI)
library(rspiro)
library(childsds)
library(arsenal)
library(lmerTest)
library(performance)
library(GLMMadaptive)
library(tidyverse)
library(ggstats)
library(corrplot)
library(patchwork)
library(gt)
library(ggbreak)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Lilah Melzer/Lilah Cough Monitor",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Lilah Melzer/Lilah Cough Monitor"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Documents/GitHub",
  "Linux" = "/home/timvigers/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

  labels = c("CF", "CF", "Healthy Control", "Healthy Control")

@@

df$Diagnosis <- relevel(df$Diagnosis, ref = "Healthy Control")

@@

    cftr_mod, cftr_mod_type, mucolytic:antih, pulm_ex, contains("micro_type___")

@@

```{r results='asis'}
#| label: tbl-mean-demo
#| tbl-cap: Participant characteristics
t1 <- tableby(
  Diagnosis ~ nih_sex + age_at_consent + height_perc + weight_perc +
    nih_race + nih_ethnicity + cftr_mod + cftr_mod_type + Genotype + pre_fev_pp +
    total_sleep + pulm_ex + micro_type___1 + micro_type___2 + micro_type___3 +
    micro_type___4,
  data = demo, control = mycontrols
)
summary(t1, pfootnote = T)
```

@@

```{r results='asis'}
per_person <- agg_counts %>%
  group_by(sid, Diagnosis) %>%
  summarise(
    `Nights` = n(),
    `Nights With Data` = sum(!is.na(Coughs)),
    `Prop. Nights With Data` = sum(!is.na(Coughs)) / n(),
    `Median Coughs per Night` = median(Coughs, na.rm = T),
    `Prop. Nights With Cough` = sum(Coughs > 0, na.rm = T) / n(),
    .groups = "drop"
  )
# Exclude the people with 0 nights of data
per_person <- per_person %>% filter(`Nights With Data` != 0)
t2 <- tableby(
  Diagnosis ~ `Nights` + `Nights With Data` + `Prop. Nights With Data` +
    `Prop. Nights With Cough` + `Median Coughs per Night`,
  data = per_person, control = mycontrols
)
summary(t2, pfootnote = T)
```

@@

```{r}
#| label: fig-coughs-dist
#| fig-cap: Histogram of Coughs Across All Nights
#| warning: false
hist1 <- ggplot(
  agg_counts[agg_counts$Diagnosis == "Healthy Control", ],
  aes(x = Coughs)
) +
  geom_histogram(binwidth = 1) +
  ylab("Nights") +
  ggtitle("Healthy Control") +
  theme_bw() +
  scale_x_break(c(80, 80), ticklabels = 80, space = 0, scales = 0.0000001)
hist2 <- ggplot(agg_counts[agg_counts$Diagnosis == "CF", ], aes(x = Coughs)) +
  geom_histogram(binwidth = 1) +
  ylab("Nights") +
  ggtitle("CF") +
  theme_bw() +
  scale_x_break(c(85, 260), ticklabels = c(260, 269), scales = 0.25)
hist <- hist1 + hist2 + plot_layout(guides = "collect")
ggsave(paste0(home_dir, "/Dissemination/Figures/histogram.png"),
  plot = hist, width = 9, height = 6, units = "in"
)
```

@@

```{r}
#| label: fig-coughs-by-night
#| fig-cap: Coughs by nights from start
#| warning: false
fig2 <- ggplot(agg_counts[!is.na(agg_counts$Coughs), ], aes(
  x = Night, y = Coughs, group = factor(sid)
)) +
  geom_point(alpha = 0.1) +
  geom_line(alpha = 0.2, aes(linetype = Diagnosis)) +
  geom_smooth(color = "black", aes(group = Diagnosis, linetype = Diagnosis)) +
  annotate("rect", xmin = 1, xmax = 29, ymin = 1, ymax = 275, alpha = 0.1) +
  annotate("rect", xmin = 41, xmax = 76, ymin = 1, ymax = 275, alpha = 0.1) +
  ylim(c(1, 275)) +
  xlim(c(0, 100)) +
  theme_bw() +
  scale_y_break(c(80, 265), scales = 0.25, ticklabels = c(265, 270, 275))
fig2
ggsave(paste0(home_dir, "/Dissemination/Figures/fig2.png"),
  plot = fig2, width = 9, height = 6, units = "in"
)
```
